Supplementary Material

### 2024-04-29 R/analysis-code/statistical-analysis.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

---
title: "statistical-analysis.qmd"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

>>>>

---
title: "statistical-analysis"
format: html
editor: visual
---

#Exploratory Data Analysis: Project Part 2

Kevin Kosewick & Emma Hardin-Parker

```{r, message=FALSE, warning=FALSE}
# Loading Necessary Packages

library(broom) 
library(here) 
library(tidymodels)
library(tidyverse)
library(corrplot)
library(ranger)
library(vip)
library(ggplot2)
library(flextable)
library(gt)
```

# Fitting Models and Statistical Analysis

#### Fitting a model using Quality of Sleep as the outcome and Physical Activity Level as a predictor

```{r}
#Load the data

sleepdata<- readRDS(here("data","processed-data","sleepdataprocessed.rds"))

lmfit_quality_activity <- lm(Quality.of.Sleep ~ Physical.Activity.Level, sleepdata)

# Placing results from lmfit_quality_activty  into a data frame with the tidy function

lmtable_quality_activity <- broom::tidy(lmfit_quality_activity)

# Viewing the results from the first model fit

print(lmtable_quality_activity)

# Saving the lmfit_quality_activity results table

lmtable_quality_activity = here("results", "tables", "lmfit1table.rds")

saveRDS(lmtable_quality_activity, file = lmtable_quality_activity)

```

The intercept value indicates that Quality of Sleep will be 6.66 if physical activity is at 0. This has a relatively low standard error and high significance. The coefficient for our variable indicates that as Physical activity level increases by one unit, Quality of Sleep will increase by 0.0109. Our t-statistic is indicated as significant by the p-value but is much lower than the intercept's. This means that Physical Activity Level does have a measurable impact on Quality of Sleep, but it's relatively small.

####Fitting a model using Quality of Sleep as the outcome and Sleep Duration as a predictor
```{r}
lmfit_quality_duration <- lm(Quality.of.Sleep ~ Sleep.Duration, sleepdata)

# Placing results from lmfit_quality_duration into a data frame with the tidy function

lmtable_quality_duration <- broom::tidy(lmfit_quality_duration)

#Viewing the results from the second model fit

print(lmtable_quality_duration)

# Saving the lmfit_quality_duration results table

lmtable_quality_duration = here("results", "tables", "lmfit2table.rds")

saveRDS(lmtable_quality_duration, file = lmtable_quality_duration)
```

For our sleep duration model, we can see that our intercept's negative value indicates that Quality of Sleep would be very poor if individuals got no sleep. This is rational and is supported by a strong p-value and a decent t-statistic, although the standard error is a bit high.

For our variable coefficient, we can see that as sleep duration increases by one unit, Quality of sleep also increases by about 1. This has a very strong p-value and t-statistic, indicating a strong relationship between this predictor and our outcome of interest.

#### Fitting a model using Quality of Sleep as the outcome and Sleep Duration and Occupation as predictors
```{r}
lmfit_quality_duration_occupation <- lm(Quality.of.Sleep ~ Sleep.Duration + Occupation, sleepdata)

# Placing results from lmfit_quality_duration_occupation into a data frame with the tidy function

lmtable_quality_duration_occupation <- broom::tidy(lmfit_quality_duration_occupation)

#Viewing the results from the third model fit

print(lmtable_quality_duration_occupation)

# Saving the lmfit_quality_duration_occupation results table

lmtable_quality_duration_occupation = here("results", "tables", "lmfit3table.rds")

saveRDS(lmtable_quality_duration_occupation, file = lmtable_quality_duration_occupation)
```

Since Occupation is a categorical variable, interpreting our intercept value is tricky. It's hard to get good information from it since we don't have any data from unemployed people. However, the coefficients for our predictors show a variety of interesting relationships.

Sleep Duration shows a similar relationship to our bivariate model we ran earlier. Sales representatives have the worst sleep quality by far with a strong t-statistic and p-value. Scientists, doctors, and sales persons are all about the next worst (with strong t-stats and p-values).

The occupations that seem to have the best sleep are teachers, lawyers, engineers and nurses. This is somewhat surprising since these are all demanding jobs that can have decently stressful work environments.

####Fitting a model using Sleep Quality as the outcome and BMI as a predictor.

```{r}
lmfit_quality_BMI <- lm(Quality.of.Sleep~ BMI.Category, sleepdata)

#Placing results from lmfit_quality_BMI into a data frame with the tidy function

lmtable_quality_BMI <- broom::tidy(lmfit_quality_BMI)

#Viewing the results from the fourth model fit

print(lmtable_quality_BMI)

# Saving the lmfit_quality_BMI results table

lmtable_quality_BMI = here("results", "tables", "lmfit4table.rds")

saveRDS(lmtable_quality_BMI, file = lmtable_quality_BMI)

```

We can see that Obese individuals have a stronger negative relationship with Sleep Quality than Overweight individuals. The lower t-statistics and p-values are likely due to fewer observations for Obese individuals in the dataset. However, it is all still statistically significant and has reasonable standard errors.

####Fitting a model using Sleep Quality as the outcome and Blood Pressure category as a predictor.
```{r}
lmfit_quality_BP <- lm(Quality.of.Sleep~ cat_bp, sleepdata)

#Placing results from lmfit_quality_BP into a data frame with the tidy function

lmtable_quality_BP <- broom::tidy(lmfit_quality_BP)

#Viewing the results from the fifth model fit

print(lmtable_quality_BP)

# Saving the lmfit_quality_BP results table

lmtable_quality_BP = here("results", "tables", "lmfit5table.rds")

saveRDS(lmtable_quality_BP, file = lmtable_quality_BP)
```

We have some interesting results for this one. Our values aren't statistically significant according to p-values, but we can see that Stage 1 Hypertension seems to have the biggest negative impact on Quality of Sleep. This is surprising since we would expect more severe hypertension to result in worse sleep, but according to our model, Stage 2 hypertension has less impact than Stage 2.

Normal blood pressure seems to have a small positive impact on sleep quality according to our model. All of these predictors seem to have a relatively small impact as they impact the sleep quality scale by less than one whole unit. Given our p-values and low t-statistics, however, this model should be taken with a grain of salt.

####Fitting a model using Sleep Quality as the outcome and Stress Level as a predictor.
```{r}
lmfit_quality_stress <- lm(Quality.of.Sleep~ Stress.Level, sleepdata)

#Placing results from lmfit_quality_stress into a data frame with the tidy function

lmtable_quality_stress <- broom::tidy(lmfit_quality_stress)

#Viewing the results from the sixth model fit

print(lmtable_quality_stress)

# Saving the lmfit_quality_stress results table

lmtable_quality_stress = here("results", "tables", "lmfit6table.rds")

saveRDS(lmtable_quality_stress, file = lmtable_quality_stress)
```

According to our model, Stress has a strongly defined relationship with sleep quality. We can see that a value of 0 on the stress score leads to a 10 unit increase in quality of sleep.

The Stress Level coefficient indicates that as stress increases by one unit, sleep quality decreases by about 0.6 of a unit. The t-statistics and p-values for the values in this model are extremely strong, indicating a well-defined relationship between stress and sleep quality.

####Fitting a model using Sleep Quality as the outcome and Age and Gender as predictors.
```{r}
lmfit_quality_gender_age <- lm(Quality.of.Sleep~ Gender + Age, sleepdata)

#Placing results from lmfit_quality_gender_age into a data frame with the tidy function

lmtable_quality_gender_age <- broom::tidy(lmfit_quality_gender_age)

#Viewing the results from the seventh model fit

print(lmtable_quality_gender_age)

# Saving the lmfit_quality_gender_age results table

lmtable_quality_gender_age = here("results", "tables", "lmfit7table.rds")

saveRDS(lmtable_quality_gender_age, file = lmtable_quality_gender_age)
```

According to our table, Males seem to have lower sleep quality than females by about a score of .0336. However, the t-statistic is very low and the p-value is very poor. Gender doesn't seem to be a good predictor of sleep quality.

When age is factored with gender, we can see that quality of sleep seems to increase slightly along with age. The t-statistic and p-value are both strong for this predictor; it may be worth doing a simple bivariate analysis of age.

####Fitting a model using Sleep Quality as the outcome and Age as a predictor.
```{r}
lmfit_quality_age <- lm(Quality.of.Sleep~ Age, sleepdata)

#Placing results from lmfit_quality_age into a data frame with the tidy function

lmtable_quality_age <- broom::tidy(lmfit_quality_age)

#Viewing the results from the eighth model fit

print(lmtable_quality_age)

# Saving the lmfit_quality_age results table

lmtable_quality_age = here("results", "tables", "lmfit8table.rds")

saveRDS(lmtable_quality_age, file = lmtable_quality_age)

```

Our model shows a strong relationship between age and quality of sleep. As Age increases, sleep quality seems to improve. Each year increase in age is predicted to have a 0.065 increase in sleep quality score.

The intercept here is also interesting as an age of "0" is predicted to have a sleep score of 4.57, which is relatively poor. The t-statistics and p-values for these are both strong.

####Fitting a model using Sleep Quality as the outcome and a newly grouped Occupatiin variable as the predictor.

Now we'll create a new variable that groups the different occupations to make analysis more succinct.
```{r}
# Create a new column

sleepdata$Occupation_Group <- as.character(sleepdata$Occupation)

# Map occupations to broader categories

sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Nurse', 'Doctor')] <- 'Healthcare'

sleepdata$Occupation_Group[sleepdata$Occupation == 'Teacher'] <- 'Education'

sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Software Engineer', 'Engineer')] <- 'Engineering'

sleepdata$Occupation_Group[sleepdata$Occupation %in% c('Accountant', 'Salesperson', 'Sales Representative', 'Manager')] <- 'Business/Finance'

sleepdata$Occupation_Group[sleepdata$Occupation == 'Scientist'] <- 'Science'

# Convert the new column to a factor

sleepdata$Occupation_Group <- as.factor(sleepdata$Occupation_Group)

## This should help with analysis of this variable. We'll fit this to a model now to test it.

lmfit_quality_occupation <- lm(Quality.of.Sleep~ Occupation_Group, sleepdata)

#Placing results from lmfit_quality_occupation into a data frame with the tidy function

lmtable_quality_occupation <- broom::tidy(lmfit_quality_occupation)

#Viewing the results from the occupation model fit

print(lmtable_quality_occupation)

# Saving the lmfit_quality_gender_age results table

lmtable_quality_occupation = here("results", "tables", "lmfit9table.rds")

saveRDS(lmtable_quality_occupation, file = lmtable_quality_occupation)

```

Looking at our newly grouped occupations, we can see that our intercept has a high t-statistic and significance; this indicates that our reference category of Business/Finance jobs have a sleep quality score of about 7, which is pretty good.

Teaching and healthcare professions don't seem to have a significant impact on an individual's sleep quality score. They have low t-statistics and large p-values.

Engineers seem to have the best sleep quality score; they're predicted to have a score of about 8.3 (with strong p-values and t-statistics). Scientists seem to have the worst score; they're predicted to have 2 points lower sleep quality than our intercept. The t-statistic and p-value are both strong for this group.

Lawyers fall in the middle, having a decent increase in sleep quality compared to the intercept and solid p-values and t-statistics to support this.

Now we'll create a table that combines all of these results to put in the manuscript.

```{r}
# Create a list of all models

models <- list(

PhysicalActivity = lmfit_quality_activity,

Occupation = lmfit_quality_occupation,

Age = lmfit_quality_age,

StressLevel = lmfit_quality_stress,

Hypertension = lmfit_quality_BP,

SleepDuration = lmfit_quality_duration,

BMI = lmfit_quality_BMI,

Gender_and_Age = lmfit_quality_gender_age

)

# Create an empty data frame to store results

all_results <- data.frame()

# Loop through each model, tidy the results, and bind to all_results

for(i in 1:length(models)) {

model_name <- names(models)[i]

model_results <- broom::tidy(models[[i]])

model_results$model <- model_name # add a column to identify the model

all_results <- rbind(all_results, model_results)

}

# Now all_results contains the combined results of all models

print(all_results)

# Save the combined results table

all_results_path = here("results", "tables", "all_results.rds")

saveRDS(all_results, file = all_results_path)

# Transform p-values to scientific notation to make table cleaner to read

all_results$p.value <- format(all_results$p.value, scientific = TRUE, digits = 2)

# Create a flextable

ft <- flextable::flextable(all_results)

# Print the table

print(ft)

# Save the flextable as an image

flextable::save_as_image(ft, path = here("results", "tables", "table.png"))

# Now we'll remove unnecessary variables (e.g. ones we've feature engineered into more useful ones) to make our analyses easier.

sleepdatafinal <- subset(sleepdata, select = -c(Occupation, systolic, diastolic, Daily.Steps, Physical.Activity.Level, Heart.Rate, AgeGroup))
```

# Creating a collinearity plot

We'll make sure there isn't collinearity between our continuous variables before we fit more complex models.
```{r}
# Select the variables

continuous_vars <- sleepdatafinal[, c("Age", "Sleep.Duration", "Quality.of.Sleep", "Stress.Level")]

# Compute correlation matrix

correlation_matrix <- cor(continuous_vars)

# Create a pairwise correlation plot (using corrplot)

correlation <- corrplot(correlation_matrix, method = "circle")

#Save the correlation plot

correlation = here("results", "figures", "correlation.rds")

saveRDS(correlation, file = correlation)
```

It seems that stress level and quality of sleep are at an absolute value of 0.9 or higher on the correlation scale. Sleep Duration and stress level also seem closely tied. Stress level and quality of sleep are both subjective scales given by an individual; this is interesting to note.

We're definitely still interested in stress as a predictor as previous work has clearly identified it as a key factor in sleep quality. We'll keep it in our analysis, but it seems good to note that it has a very linear relationship with sleep quality.

# Fitting a Random Forest model using cross validation

First, we'll create a random seed to aid in reproducibility.
```{r}
#Set the seed
rngseed=1234

set.seed(rngseed)
```

We'll now train a Random Forest model using CV with no train/test split (as the data has less than 400 observations and many of the values are rare/unique for some columns). We'll follow the tidymodels framework.

```{r}
# Random Forest model specification; include the mtry and min_n parameters

model_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 300) %>%

set_engine("ranger", importance = 'impurity', seed = rngseed) %>%

set_mode("regression")

# Specify the recipe

data_recipe <- recipe(Quality.of.Sleep ~ ., data = sleepdatafinal)

# Create Random Forest workflow

data_workflow <- workflow() %>%

add_model(model_spec) %>%

add_recipe(data_recipe)

# Create resamples using 5-fold cross-validation, 5 times repeated

resamples <- vfold_cv(sleepdatafinal, v = 5, repeats = 5)

# Define the grid of parameters

mtry_param <- mtry(range = c(1, 7))

min_n_param <- min_n(range = c(1, 21))

rf_grid <- grid_regular(mtry_param, min_n_param, levels = 7)

# Tune the Random Forest model

tune_results_rf <- tune_grid(

data_workflow,

resamples = resamples,

grid = rf_grid

)

# Print RF results

print(tune_results_rf)

#Plot them too

autoplot(tune_results_rf)
```

We've successfully fit our RF model; now, we can pick the different pieces apart and see what predictors seem the most important and which model is best.

####Best RMSE model

```{r}
best_params <- select_best(tune_results_rf, metric = "rmse")

#Variable importance info

# Finalize the workflow with the best parameters

final_wf <- finalize_workflow(data_workflow, best_params)

# Fit the finalized workflow to the entire data

final_fit <- fit(final_wf, data = sleepdatafinal)

# Extract the fitted model

fitted_model <- extract_fit_parsnip(final_fit)

# Create a variable importance plot

vip(fitted_model$fit)

```

We can see that sleep duration, stress level, and Age seem to have the most importance for our model.

PhysicalActivityLevel doesn't seem to have much of an impact on Quality of Sleep, as it was excluded from the variable importance plot altogether. These results are pretty in line with what we discovered with our EDA and our simple linear models, but now we have a ranking given by this plot.

Now, we'll see how well this model predicts on new data.

```{r}
# Make predictions

predictions <- predict(final_fit, new_data = sleepdatafinal)

# Add the predictions to dataframe

sleepdatafinal$Predicted <- predictions$.pred

# Create an observed vs. predicted plot

ggplot(sleepdatafinal, aes(x = Quality.of.Sleep, y = Predicted)) +

geom_point() +

geom_abline(color = "red") +

labs(x = "Observed", y = "Predicted", title = "Observed vs. Predicted Plot") +

theme_minimal()

#Now we'll calculate the RMSE and R squared values of our "best" set of parameters

# Calculate residuals

residuals <- sleepdatafinal$Quality.of.Sleep - predictions$.pred

# Calculate RMSE

rf_rmse <- sqrt(mean(residuals^2))

# Calculate R squared

sst <- sum((sleepdatafinal$Quality.of.Sleep - mean(sleepdatafinal$Quality.of.Sleep))^2)

ssr <- sum(residuals^2)

rf_r_squared <- 1 - (ssr / sst)

#print results

print(rf_rmse)

print(rf_r_squared)
```

This seems like a very solid fit; not too close to where overfitting is obvious, but not too far to where the predictions aren't useful either. The RMSE is extremely low (.059) and the R squared is extremely high (0.997). This is a good indicator that our selected model is powerful.

#### Random Forest Model Evaluation

Now, we'll make a "null" model and compare its results to that of our RF in order to make sure that the RF is useful.
```{r}
# Specify the null model

null_mod <- null_model() %>%

set_engine("parsnip") %>%

set_mode("regression")

# Fit the null model to the data

null_fit <- fit(null_mod, Quality.of.Sleep ~ 1, data = sleepdatafinal)

# Make predictions

null_predictions <- predict(null_fit, new_data = sleepdatafinal)

# Bind the predictions with the actual values

data_with_predictions <- bind_cols(sleepdatafinal, null_predictions)

# Compute RMSE

null_rmse <- rmse(data_with_predictions, truth = Quality.of.Sleep, estimate = .pred)

# Print the RMSE

print(null_rmse)

```

This RMSE is a good bit higher than the 0.2-0.32 range of all of our Random Forest models. This is a good sign that the RF model performs well on the data.

```{r}
# Compute R-squared

r_squared <- rsq(data_with_predictions, truth = Quality.of.Sleep, estimate = .pred)

# Print the R-squared

print(r_squared)

# Create a predicted v observed plot

ggplot(data_with_predictions, aes(x = Quality.of.Sleep, y = .pred)) +

geom_point() +

geom_abline(color = "red") +

labs(x = "Observed", y = "Predicted", title = "Observed vs Predicted Null Values") +

theme_minimal()

```

Since there is no variation between the model predictions and the observed values in a null model, the R squared is not possible to generate. The observed vs predicted plot is also as expected; a straight line of predictions that all match the average sleep quality score.

Now we'll make a table for the manuscript comparing the metrics we evaluated.

```{r, message=FALSE}
#extract the metric values for the null model from the objects they're stored in

null_rmse_value <- null_rmse$.estimate

null_r_squared_value <- r_squared$.estimate

# Create a data frame

results <- data.frame(

Model = c("Random Forest", "Null Model"),

`R-squared` = c(rf_r_squared, null_r_squared_value),

RMSE = c(rf_rmse, null_rmse_value)

)

# Create the table

gt_table <- gt(results)

# Print the table

print(gt_table)
```

### 2024-04-29 products/manuscript/Manuscript.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-18-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/references_sleepdata.bib
editor: 
  markdown: 
    wrap: 72
---

@@

For our project, we obtained a dataset from
[Kaggle](https://www.kaggle.com/datasets) containing 373 observations of
13 variables from a study on different factors that may affect sleep
quality. Variables of interest from the study included physical activity
level, age, sex, occupation, stress, gender, BMI, blood pressure, and
sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/)
(version 4.3.1), we used a combination of packages to create models
suitable for answering our primary question: which variables are most
important for determining an individual's sleep quality and what effect
do they have? After conducting an exploratory data analysis (EDA) we
hypothesized that stress, sleep duration, and occupation would have the
largest impact on sleep quality. We believed that stress and sleep
quality would be negatively correlated, sleep duration and quality would
be positively correlated, and that jobs with longer or more variable
hours would negatively affect sleep quality.

@@

Engaging in physical activity has been suggested to improve both sleep
quality and sleep duration. Regular moderate to intense physical
activity stimulates an increase of melatonin production, a hormone that
regulates sleep-wake cycles (Alnawwar et. al, 2023). Additionally,
moderate to intense exercise has shown to have a positive effect on
those with insomnia (Alnawwar et. al, 2023). Physical activity also
reduces stress and improves both sleep quality and sleep efficiency
(Alnawwar et. al, 2023). However, studies also show that frequent
high-intensity workouts could lead to poorer sleep quality among people
with and without insomnia (Alnawwar et. al, 2023). Studies indicate that
consistent moderate exercise three to four times a week (approximately
150 minutes total) is the optimal weekly regimen to improve sleep
quality and reduce stress. We expect to see similar trends with our
data, where increased stress levels negatively impact sleep quality and
sleep duration, while noticing a positive association between increased
physical activity levels and sleep quality.

@@

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. There were no missing values
or erroneous variables in the dataset that we needed to remove. The
"Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by \[10,000 steps.org\]
(https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:\~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.

@@

The main limitations of our models stemmed from the data itself. As
stated in our introduction section, we have been unable to identify
exactly where the data were collected and how it was gathered. The
values for stress level and sleep duration were also based on a
subjective rating given by each participant in the study; this could
introduce individual bias to the data and prevent an objective
evaluation of the effect of these variables. An inherent limitation in
RF models is difficulty of interpretation; we attempted to minimize this
by creating easy to read tables and graphs for the results. Future
analyses would benefit from more detailed meta data and a better
understanding of the experimental design that generated our
observations.

>>>>

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/references_sleepdata.bib
editor: 
  markdown: 
    wrap: 72
---

@@

For our project, we obtained a dataset from
Kaggle.com containing 373 observations of
13 variables from a study on different factors that may affect sleep
quality. Variables of interest from the study included physical activity
level, age, sex, occupation, stress, gender, BMI, blood pressure, and
sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/)
(version 4.3.1), we used a combination of packages to create models
suitable for answering our primary question: which variables are most
important for determining an individual's sleep quality and what effect
do they have? After conducting an exploratory data analysis (EDA) we
hypothesized that stress, sleep duration, and occupation would have the
largest impact on sleep quality. We believed that stress and sleep
quality would be negatively correlated, sleep duration and quality would
be positively correlated, and that jobs with longer or more variable
hours would negatively affect sleep quality.

@@

Engaging in physical activity has been suggested to improve both sleep
quality and sleep duration. Regular moderate to intense physical
activity stimulates an increase of melatonin production, a hormone that
regulates sleep-wake cycles (Alnawwar et. al, 2023). Additionally,
moderate to intense exercise has shown to have a positive effect on
those with insomnia (Alnawwar et. al, 2023). Physical activity also
reduces stress and improves both sleep quality and sleep efficiency
(Alnawwar et. al, 2023). However, studies also show that frequent
high-intensity workouts could lead to poorer sleep quality among people
with and without insomnia (Alnawwar et. al, 2023). Studies indicate that
consistent moderate exercise three to four times a week (approximately
150 minutes total) is the optimal weekly regimen to improve sleep
quality and reduce stress. We expect to see similar trends with our data
where increased stress levels negatively impact sleep quality. We expect to see
a positive association between increasing physical activity levels and
sleep quality until the point of "too much" activity.

@@

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. We have added this codebook to
the supplementary materials file for ease of access but you can also
find it on the Kaggle page for this dataset. There were no missing
values or erroneous variables in the dataset that we needed to remove.
The "Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.

@@

The main limitations of our models stemmed from the data itself. As
stated in our introduction section, we have been unable to identify
exactly where the data were collected and how it was gathered. The
values for stress level and sleep duration were also based on a
subjective rating given by each participant in the study; this could
introduce individual bias to the data and prevent an objective
evaluation of the effect of these variables. An inherent limitation in
RF models is difficulty of interpretation; we attempted to minimize this
by creating easy to read tables and graphs for the results.
We chose not to split our data into a train/test split due to the small number of observations and presence of rare values in our data. We believe that fitting using CV minimized the downsides of this. Future
analyses would benefit from more detailed meta data and a better
understanding of the experimental design that generated our
observations.

### 2024-04-29 products/manuscript/supplement/Supplementary-Material.qmd
Final Part Revisions
Made statistical analysis file a .qmd and have cleaned some issues pointed out in the project reviews.

*This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.*

*For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).*

*I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.*

### I HAVE STRUCTURED THIS FILE. NEED TO WORK ON PULLING CERTAIN FIGURES, PLOTS, TABLES, ETC. ###

@@

@@

*Example additional result*

*@tbl-resulttable1 shows an additional table summarizing a model fit.*

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

*# Discussion*

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

# References

>>>>

@@

The following information is the codebook for this dataset:

@@
